A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-09-24 paper/paper.tex
paper: polish sec:ett a bit

The definition is dispatched over tables \ref{tab:coind}, \ref{tab:core}, \ref{tab:frames} and \ref{tab:faces}. The extra table~\ref{tab:coh} shows the proof of a coherence condition used in \ref{tab:faces}. Table \ref{tab:coind} describes a $\nu$-set in indexed form, as a coinductively-defined infinite sequence of types representing the limit of $n$-truncated $\nu$-sets. The definition of a $n$-truncated $\nu$-set is in turn described in \ref{tab:core}. In the infinite sequence of types representing a $\nu$-set, the $n$-th component is a type dependent over a $\fullframe$. The type $\fullframe$ is recursively defined in \ref{tab:frames}, using the auxiliary definitions of $\layer$ and $\painting$. The type $\layer$ and $\painting$ are dependent over $\restrf$, and these restrictions are defined on \ref{tab:faces}. These restrictions are defined using auxiliary definitions of $\restrl$ and $\restrc$.

Notably, the definition of $\restrl$ relies on a definitional equality expressing the commutation of the composition of $\restrf$. This commutation is not provable by computation so we have to prove it propositionally before using the reflection rule. Proving this itself requires an induction on the dimension, and on the structure of $\framep$, $\layer$, and $\painting$. This is what $\cohf$ proves, as shown in the table \ref{tab:coh}, using auxiliary definitions $\cohl$ and $\cohc$. Even though it looks independent of the other tables, $\cohf$ has to be proved mutually with the definitions of $\framep$, $\layer$, $\painting$, and their corresponding restrictions. More precisely, for a fixed $n$, the block of $\framep$, $\restrf$, and $\cohf$ has to be mutually defined by induction on $p$. Also, each of $\painting$, $\restrc$, and $\cohc$ is built by induction from $p$ to $n$. The $\painting$ block at $n$ relies on the $\framep$ block at $n$, but, the converse dependency is only on lower $n$, so this is well-founded. Note that $\layer$, $\restrl$ and $\cohl$ are just abbreviations. We leave however implicit the exact way this mutual recursion can be formalized at this stage of the paper.

Most components of the construction takes inequality constraints as parameters, and we have left implicit that they are satisfied in the tables. Comparison of inequality proofs depends on the definition of inequality, which we leave implicit, only assuming that we can pick a definition for which proofs of $m \leq n$ are provably unique and thus definitionally equal by the reflection rule.

Note that for a fixed constant $n$, the coherence conditions evaluate to a reflexivity proof, so that the construction evaluates to an effective sequence of types of iterated relations not mentioning them anymore.

@@

\subsection{From extensional to intentional type theory: explicitation of the equality proofs\label{sec:itt}}
\renewcommandx{\cohf}{\coh{frame}[][][][][][][][]}

@@

The way recursion is implemented is still left implicit at this stage. Also, depending on how inequality on natural numbers is defined, different equalities may hold or not definitionally, such as transitivity of inequality or the computational properties of induction over inequality proofs. Details are given in \ref{sec:mech}.

>>>>

The definition is dispatched over tables \ref{tab:coind}, \ref{tab:core}, \ref{tab:frames} and \ref{tab:faces}. The extra table~\ref{tab:coh} shows the proof of a coherence condition used in \ref{tab:faces}. Table \ref{tab:coind} describes a $\nu$-set in indexed form, as a coinductively-defined infinite sequence of types representing the limit of $n$-truncated $\nu$-sets. The definition of a $n$-truncated $\nu$-set is in turn described in \ref{tab:core}. In the infinite sequence of types representing a $\nu$-set, the $n$-th component is a type dependent over a $\fullframe$. The type $\fullframe$ represents a matching object. It is recursively defined in \ref{tab:frames}, using the auxiliary definitions of $\framep$, $\layer$ and $\painting$. The type $\layer$ relies on an operator of frame restriction $\restrf$ which is defined in \ref{tab:faces}. This restriction operator is defined using the auxiliary definitions of $\restrl$ and $\restrc$.

Notably, the definition of $\restrl$ relies on an equality expressing the commutation of the composition of $\restrf$. This commutation is not provable by computation so we have to prove it propositionally before using the reflection rule. Proving this itself requires an induction on the dimension, and on the structure of $\framep$, $\layer$, and $\painting$. This is what $\cohf$ proves, as shown in the table \ref{tab:coh}, using auxiliary definitions $\cohl$ and $\cohc$. Even though it looks independent of the definitions from the the other tables, $\cohf$ has to be proved mutually with the definitions of $\framep$, $\layer$, $\painting$, and their corresponding restrictions. More precisely, for a fixed $n$, the block of $\framep$, $\restrf$, and $\cohf$ has to be mutually defined by induction on $p$. Also, each of $\painting$, $\restrc$, and $\cohc$ is built by induction from $p$ to $n$. The $\painting$ block at $n$ relies on the $\framep$ block at $n$, but, the converse dependency is only on lower $n$, so this is well-founded. Note that $\layer$, $\restrl$ and $\cohl$ are just abbreviations. The exact way this mutual recursion is eventually formalized is explained in section~\ref{sec:wf}.

Most components of the construction takes inequality constraints as parameters, and we have left implicit that they are satisfied in the tables. Comparison of inequality proofs depends on the definition of inequality, which we leave implicit, only assuming that we can pick a definition for which proofs of $m \leq n$ are provably unique and thus definitionally equal by the reflection rule (formal details are given in section~\ref{sec:le}).

Note that for a fixed constant $n$, relying on the evaluation rules of type theory, the coherence conditions degenerate to a reflexivity proof, so that the construction builds an effective sequence of types not mentioning coherences anymore.

@@

\subsection{From extensional to intentional type theory: explication of the equality proofs\label{sec:itt}}
\renewcommandx{\cohf}{\coh{frame}[][][][][][][][]}

@@

The way recursion is implemented is still left implicit at this stage. Also, depending on how inequality on natural numbers is defined, different equalities may hold or not definitionally, such as transitivity of inequality or the computational properties of induction over inequality proofs.
